---
title             : "The categorization of L3 sounds in Spanish/English bilinguals at the very initial stage of L3 acquisition"
shorttitle        : "The categorization of L3 sounds in Spanish/English bilinguals"

author: 
  - name          : "Kyle Parrish"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "15 Seminary Place, New Brunswick, NJ 08904"
    email         : "kcp74@scarletmail.rutgers.edu"

affiliation:
  - id            : "1"
    institution   : "Rutgers University"




keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_docx
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

Much recent research in third language acquisition (L3A) has been primarily concerned with the nature of cross-linguistic influence (CLI) at the very initial state of L3 acquisition and has been focused on morphosyntax. Bilinguals have, by definition, 2 linguistic systems available for transfer when they begin to learn a third language, and the precise role of these languages in L3A has been debated. For instance, some scholars have suggested that the second language learned in late bilinguals (those who began the acquisition of the L2 after childhood) blocks access to the L1, and is the primary influence to the third language (The L2 Status Factor; @bardel_role_2007. Others posit that language typology plays a decisive role in determining whether the L1 or the L2 is the primary influence on the third language [@baauw_cognitive_2013; @rothman_l3_2011; @rothman_linguistic_2015; @rothman_typological_2010]. In this view, the individual L3 learners’ subconscious impression of cross-linguistic similarity between the languages that they know in comparison to the language that they are learning determines which language primarily influences their L3. Contrary to both the typological and L2 status views, which predict that only a single language transfers to the L3, other models suggest that cross-linguistic influence may occur on a property by property basis [@slabakova_scalpel_2017;@westergaard_crosslinguistic_2017]. 
Importantly, these models have largely received support from empirical studies in morphosyntax, and thus make predictions of morphosyntactic transfer into a third language. It is unclear whether crosslinguistic interactions of phonetics can also be predicted by the L3 morphosyntax models, since they are constrained to a larger degree in their examination of potential gradient effects in L3 learners. For instance, two language word orders (SOV vs SVO) can directly and categorically be observed experimentally in L3 learners, in which either L1 word order or L2 word order preference is observed in a single structure of the L3. On the other hand, and given the lack of gradience in the realization of syntactic transfer (particularly in a single observation), L3 phonetics allow for a more precise examination of the role of the L1 and the L2 to be examined in the production or perception of an acoustic signal. For example, measurements of the acoustic realizations of L3 productions of segment, such as voice onset time (VOT) in word-initial stop consonants, may provide evidence not only for L1, L2 or property-by-property transfer, but a gradient value, which may also vary by property, and be influenced in tandem by both existing languages in the bilingual mind. This information, together with findings in morphosyntax, can paint a clearer picture of the factors which drive CLI and how CLI interacts across phonology and morphosyntax. 
	The present study contributes to research on the role of the L1 and the L2 at the very beginning of L3 acquisition by examining the potential gradient effects of both languages on the perception of an unknown third language by Spanish/English bilinguals. Additionally, the present study’s use of four groups experimentally controls for L2 status effects and typology. First, Spanish English bilingual participants were recruited in both orders of acquisition (English speakers who grew up monolingual and learned Spanish post-childhood and vice versa). These so-called mirror image groups have been used in previous studies [@puig-mayenco_systematic_2020] and avoid the confound of typological transfer and L2 status (i.e. when L2 status effects are observed, it is not simply typologically driven transfer that happens to also be the second learned language). Secondly, the present study aims to examine how Spanish-English bilinguals categorize third language sounds by observing how they categorize two continua (a VOT and F2 continuum) in two different third languages, what they believe to be Hungarian (unclear relationship with Spanish and English) or French (clear Spanish-French connection). Third, the combination of the perception of two continua in the present study, and their measurement in all three languages allows for the possibility of both gradient effects from both the L1 and the L2, a phenomenon that is much more difficult to observe in syntax, and a difference in this effect in two sound distinctions that may suggest that CLI occurs both gradiently and is structure specific. 
In sum, the present study allows for the testing of four potential outcomes in the beginning stages of L3A: typological transfer, L2 transfer, structure-by-structure transfer, or gradient transfer. The present study that phonetics are the an ideal candidate for the examination of these four possiblities, given its meaurement in continuous values.

# Literature Review

In the past ten years, the field of third language acquisition has grown considerably. L3 models of morphosyntax attempt to determine the factors which drive transfer at the beginning stages of learning a third language. Broadly, these models fall under two categories: full transfer models, which posit that one language is transferred to the L3 at the initial stages of L3A (also referred to as holistic transfer), and property-by-property models [The Linguistic Proximity Model, LPM; @westergaard_crosslinguistic_2017, the Scalpel Model, @slabakova_scalpel_2017]. The body of this literature review will first explain full transfer models of morphosyntax and their empircal evidence, followed property-by-property models. Then, studies in L3 phonology will be covered in detail. Finally, recent research in categorical perception in bilinguals will be discussed. These studies in the categorization of language specific sounds serve as the methodological basis to the present study.

## Full Transfer models of L3 Morphosyntax

### The Typological Primacy Model
The Typological Primacy Model [TPM; @baauw_cognitive_2013; @rothman_l3_2011; @rothman_linguistic_2015; @rothman_typological_2010] posits that transfer of one grammar holistically occurs in the initial stages of the L3, rather than on a structure-by-structure basis. This initial transfer is based on a subconscious “choice” or so called psycho-typology [@kellerman_now_1983] made by the parser based on the cues available in the L3 input. In particular, TPM suggests that the hierarchy for the cues used by the L3 learner to determine the typological similarity of languages (and thus which language transfers) is primarily based on cues in the L3 lexicon, followed by phonological/phonotactic information, then by functional morphology and lastly by syntactic structure. That is, in the view of the TPM, the L3 learners will first attend to cross-linguistic similarities in the lexicon to which they are exposed via L3 input to make a determination (the L1 or the L2) of which system to transfer to the L3. Importantly, this view of the TPM requires that the L3 learner is exposed to input. @rothman_linguistic_2015 explains that the “initial stages” of L3A are some 20-25 hours of instruction in a third language, since the L3 learner must be exposed to input before determining similarity between linguistic systems. In the present study, perceived typological proximity is predicted to be seen in distinct categorizations of what the participants believe to be French sounds as opposed to Hungarian sounds.

### The L2 Status Factor
The L2 Status Factor [L2SF: @bardel_role_2007; @cabrelli_amaro_l2_2012] posits that the second language learned serves as the source of transfer to a third language by default, regardless of typology. Framed within the Declarative/Procedural Model [@paradis_declarative_2009; @paradis_neurolinguistic_2004; @ullman_contributions_2004], the L2SF suggests that the cognitive similarity of the L2 and the L3 may serve as a neurolinguistic explanation for the findings that the L2 transfers by default. That is, within the DP model, adult L2 learners rely more heavily on declarative memory than procedural memory to learn the grammar of a new language, as opposed to L1 learning, which is learned in large part by children using procedural memory. Since the L3 would likely also involve a higher reliance on declarative memory, it has a higher degree of similarity to the L2, which is thought to drive L2 transfer. It is important to note that the L2 Status Factor does not make claims for simultaneous bilinguals who learn a third language later in life, or early trilinguals. The claims made by the model refer to late bilinguals, since it is thought that they rely on declarative memory to learn a language. 
However, studies involving the L2 Status Factor have found that it is possible to transfer from the L1, when L3 learners have higher metalinguistic knowledge. @falk_role_2015 examined L3 learners of a V2 language when their L1 was also a V2 language. Using a metalinguistic awareness survey, they found that only the participants who scored higher on the survey were able to transfer their L1 V2 properties to the L3 (there was a wide range of non-V2 L2s used in the study).

### Evidence for full transfer models 
Conflicting evidence has been found in support of both of these full transfer models. For instance, in a recent meta-analysis of 92 studies in L3 acquisition, @puig-mayenco_systematic_2020 determined that typology (59 out of 92) and L2 status (29 out of 92) were the best predictors of which linguistic system would serve as the basis for a third language. However, these findings were not mutually exclusive; the authors coded 25 total studies as being explained both by L2 status and typology, meaning that the results of these studies reported that the L2 transferred to the L3, but could not rule out the possibility that psycho-typological transfer could also explain the results, since the studies did not use mirror image groups (i.e. L3 groups with the same languages, but the opposite order of acquisition). An issue with this explanation, however, is the lack of variance in the studies that the authors coded as both typological and L2 transfer which is predicted to occur by the TPM. If typological transfer is individually determined after exposure to L3 input (input which likely is distinct for each L3 learner), and not necessarily a reflection of actual underlying structural similarities [@rothman_linguistic_2015], then some variance would be expected to be observed in populations in which cross-linguistic typology is unclear. In other words, one would expect that, based on the predictions of the TPM, that some learners would transfer their L1 and others would transfer their L2, particularly when there is not a clear typological similarity between the languages involved, rather than entire groups transferring solely their L1 or their L2. Only one study has reported variance in the choice of the L1 and L2 in L3 acquisition in L3 learners whose order of acquisition was the same (i.e. not mirror image groups seemingly driven by typology). In the study on L3 learners of Swedish, @falk_role_2015 found that some learners transferred their L1 and others transferred their L2 in their judgments of V2 sentences. The participants all spoke a V2 L1, and non-V2 L2, but differed in their overall metalinguistic knowledge. The study found that higher participants with higher metalinguistic knowledge were able to avoid L2 status effects, which would have resulted in non-facilitative transfer from their L2, and showed facilitative transfer from L1. Although the authors suggest that this transfer is due to metalinguistic knowledge, rather than underlying perceived psychotypology, these results could also be explained by the TPM, since psycho-typological transfer is made on an individual basis. Other studies have found that either the L1 or the L2 may transfer holistically. In a series of studies investigating the acquisition of L3 Portuguese by Spanish/English bilinguals, Rothman and colleagues found that Spanish transferred to L3 Portuguese, whether it was the L1 or L2 of the participants. Importantly, these results do not exclude the possibility that metalinguistic knowledge drives the choice of the choice of a language to holistically transfer, nor do these studies rule out the possibility of structure-by-structure transfer, since they typically focus on just one property per study. It is possible that the relatedness of language combinations such as Spanish, Brazilian Portuguese, and English and the fact that the participants in many studies are university students combine to make very salient that Spanish and Portuguese are closely related languages. Overall, the empirical support of TPM falls victim to a similar issue; the combination of English and two romance languages shows typological transfer [@puig-mayenco_systematic_2020]. The lack of support for the TPM outside of this particular language combination and the lack of variance in individual choice of L1 or L2 beg the question: could the results of the TPM be explained by top down processes, and knowledge of language similarity, and unclear relationships show L2 status effects by default?

## Hybrid Transfer models of L3A

### The Linguistic Proximity Model
Unlike the TPM and the L2SF, the Linguistic Proximity Model [@westergaard_crosslinguistic_2017] suggests that transfer occurs on a property-by-property basis, rather than holistically. The LPM suggests that there is a “full transfer possibility”, meaning that any individual structure may transfer at any time, but that it also may not. Scholars have argued that this vague prediction creates a problem in modeling L3 transfer acquisition [@bardel_l1_2020; @wrembel_multilingual_2020], since it is unclear when transfer of a particular structure occurs and when it does not. Likewise researchers have argued that, unlike the TPM and the L2SF, the LPM is not easily falsifiable [@bardel_l1_2020]. Given the lack of predictive power of the LPM, it may only receive post-hoc support. In the present study, influence from different languages in the two different continua would provide support for the LPM, but the LPM would not have predicted how and why transfer from one language affects a particular property of the L3.

## L3 phonology

### Studies in L3 phonology

Similar to studies in L3 morphosyntax, the relatively short history of L3 phonology has endeavored to identify the source and direction of cross-linguistic influence in the linguistic sound systems of multilinguals, with particular interest in the influence of the two previously acquired systems on the third language system being acquired [@cabrelli_amaro_investigating_2016]. 
Various studies have found a stronger L2 influence on L3 productions at the early stages of L3 acquisition. In a case study foundational to L3 phonology, @hammarberg_articulatory_1993 and @hammarberg_processes_2005 examined the productions of a single informant, who spoke L1 British English, L2 German, and L3 Swedish. SW, a near-native German speaker as rated by German natives, was recorded reading a story in Swedish at two different time points, first one month after moving to Sweden, and second after one year having lived there. Those Swedish recordings at both time points were then judged by native speakers of Swedish who were to judge the accent of the speaker in the story, unaware that they were hearing the same informant at two different times. Interestingly, the Swedish speakers rated the first telling of the story as heavily German accented (influence from the L2), whereas in recording 2, they rated the accent as coming from an English speaker. This initial effect has been referred to as a ‘foreign language effect’ [@meisel_transfer_1983] or, more recently, as L2 status [@bardel_role_2007; @hammarberg_processes_2005]. The findings of @hammarberg_processes_2005 suggest that this foreign language effect, or, in other words, the desire to sound less like a speaker of one’s L1, may diminish as proficiency increases and be replaced by L1 effects. Other studies have also found primarily L2 influence in L3 productions in global accent ratings [@wrembel_l2-accented_2010], VOT productions [@llama_influence_2010; @tremblay_l2_2007], vowel production [@kamiyama_acquisition_2007] vowel reduction and speech rhythm [@gut_cross-linguistic_2010].
Other findings in L3 phonology production, however, have yielded mixed results. Several studies have found that productions of values fall between L1 and L3 values, suggesting that both previously the L1 and the L2 have some influence on L3 productions. For instance, [@wrembel_vot_2014] measured VOT and aspiration in all languages of participants with two different language combinations: L1 Polish, L2 English, and L3 French; (2) L1 Polish, L2 English, and L3 German. The results showed that each language had a specific stop-value, and that the L3 VOT productions were intermediate, falling between the L1 and L2 values. Similarly, [@wrembel_cross-linguistic_2011] examined thirty-two learners of L3 French with L1 Polish and L2 English who were recorded reading lists of words in carrier phrases. Similarly to [@wrembel_vot_2014], combined transfer from the L1 and the L2 in VOT productions was found. Similar findings in VOT productions were also reported by @wunder_phonological_2010 in L3 Spanish speakers. @blank_transferencia_2009 also found intermediate values in VOT productions in vowel production of L3 English speakers who spoke L1 Brazilian Portuguese and L2 French. Other studies have found an L1 influence on production despite L3 proficiency [@wrembel_foreign_2012], or at advanced L3 learners [@llama_revisiting_2018].
	Perception studies in L3 acquisition have been much more scarce than those on production, and primarily in young participants. One of very few L3 perception studies in adults is @liu_effects_2019, which examined the perceptual boundary of a VOT continuum in trilinguals. The participants were L3 spanish speakers who spoke L2 English and L1 Chinese. Though the authors focused their analysis on regressive transfer and comparisons to all speakers, the reported boundaries in the study (n =10, 28ms for Chinese, 24.6ms for English and 23ms for Spanish) suggest that the participants were using their English boundaries for L3 Spanish categorization.  Other studies in L3 perception, which have examined young trilinguals, have found a wide range of results. For instance, @wrembel_cross-linguistic_2020 found that cross-linguistic influence is structure dependent and varies among individuals, while @balas_perception_2019 found CLI to be modulated by a complex combination of factors such as markedness of the segment under examination, proficiency in the L2 and the L3, and L1 typology. Finally, another study on L3 perception in young learners revealed that L2 predominantly influences L3 perception in a range of sounds [@wrembel_cross-linguistic_2020]
	These wide ranging results, ranging from L1 influence, L2 influence, a combined influence of L1 and L2, and a structure dependent influence of either L1 or L2 on the L3 may be the result of the wide range of realizations, or lack of homogeneity, in multilinguals. Firstly, age effects in past studies have varied considerably, in both the age of onset of the L2 and the L3. It is worth noting that many studies which report L2 status effects are in adult L3 learners, who have a late age of onset in their L2 [@hammarberg_processes_2005; @wrembel_l2-accented_2010] or this age information was not reported [@gut_cross-linguistic_2010; @kamiyama_acquisition_2007; @tremblay_l2_2007]. On the other hand, studies in which L3 learners began their acquisition of their second language at a young age (under 12) tend to show a combined effect of both the L1 and the L2 [@blank_transferencia_2009; @wrembel_cross-linguistic_2011; @wrembel_vot_2014; @wunder_phonological_2010]. Although no study has investigated age effects and L3 transfer directly, the L2 Status Factor [@bardel_role_2007] can explain these findings based on the declarative procedural model [DP Model; @paradis_declarative_2009; @paradis_neurolinguistic_2004; @ullman_contributions_2004]. Based on the proposed cognitive similarities between late-learned languages in the DP model, the L2 status factor posits that the L2 will transfer to the L3 at the initial stages of acquisition. In this view, native language grammars are known implicitly and largely subserved by the procedural memory, whereas late learned language grammars (the L2 and beyond) are thought to be more explicitly known and subserved by the declarative memory. Importantly, the L2SF does not make predictions in regard to young bilinguals who learn a third language, precisely because the DP model would predict that both languages of young bilinguals would be implicitly known and both exist in the procedural memory. This is in contrast to late bilinguals who learn a third language, whose native language grammar is procedural and whose L2 and L3 are declarative. 
	The factor of proficiency in both the L2 and the L3 is another vital aspect to language transfer, and has been subject to some controversy among scholars. In terms of proficiency in the L2, the DP model suggests that, as proficiency grows, the L2 becomes more subserved by procedural memory, and less dependent on the explicitly driven declarative memory system. This gradual shift in dependence may result in a less clear and universal choice of language system to transfer by L2 learners in the beginning stages of L3 acquisition. That is, it is possible that late bilinguals who are highly proficient in their L2 may pattern more like early bilinguals than late bilinguals who have lower L2 proficiency. Other scholars argue that L2 proficiency must be sufficiently high in order for the L2 to transfer to the L3 [@rothman_linguistic_2015]. Taken together, it would seem that a threshold for L2 proficiency is necessary in order for L2 status effects to be observed. 
L3 phonology studies have found conflicting evidence in the role of proficiency, and have focused on the L3, rather than L2 proficiency. Some scholars suggest that the L2 serves as the basis for L3 transfer at low levels of L3 proficiency as a sort of ‘coping mechanism’ [See @wrembel_search_2015 for a review]. Other studies have found that L2 accented L3 speech is present in low proficiency L3 learners but that higher L3 proficiency speakers show L1 effects in their productions of their L3 [@hammarberg_processes_2005; @wrembel_l2-accented_2010]. In sum, both L2 and L3 proficiency levels are vital and must be measured or statistically controlled to determine the precise effect of the proficiency of both languages on CLI. 


### Categorical Perception

The present study will inform models of L3 morphosyntax by examining categorical perception of bilinguals, and determining which of the language specific categorical (phonemic) boudaries Spanish/English bilinguals use to account for L3 sounds at their first exposure. Of the L3 morphosyntax models, only the TPM makes predictions in regard to phonological information. Namely, it predicts that phonological or phonotactic cues are used to determine typological proximity when lexical overlap cannot readily be detected by the parser in the L3 input [@rothman_linguistic_2015]. Since the precise nature of the use of phonological information in the TPM is unclear and untested, the present study presents the same two continua to participants in order to measure their categorization of what they believe to be English, Spanish, and a third language that they do not know. This manipulation of language mode has been carried out successfully in previous studies in both conceptual and perceptual cueing which result in language specific categorization of the same continuum. For instance, @casillas_perceptual_2018 tested a VOT continuum of pseudowords pafri-bafri (originally used in @gonzales_bafri_2013) in a 2 alternative forced choice task in learners of L2 Spanish. The stimuli differed in specific language that they were perceptually cued by a Spanish and English language specific rhotic in the second syllable. The results revealed that two language specific boundaries were found in the perception of the initial stop sound in sufficiently proficient bilinguals.
In a follow up study, @lozano-arguelles_conceptually_2020 also investigated Spanish English bilinguals’ categorization of word initial stops /b/ and /p/. Differently from @casillas_perceptual_2018 and @gonzales_bafri_2013, this study presented only the first syllable of the pafri-bafri continuum (paf-baf), and manipulated the participants’ language mode using conceptual cueing by telling participants they were listening to English or Spanish sounds, rather than perceptual cueing of the language specific rhotic sounds used by previous studies. The results were consistent with previous studies; sufficiently proficient L2 Spanish speakers show language mode effects in their categorization of /p/ and /b/ on a VOT continuum, whether this is perceptually or conceptually cued. In other words, both language specific phonetic detail within the words were sufficient to bring about that language’s categorization of the stop consonant, as well as top-down information that lead the participants to apply on of their two language specific boundaries as a function of which language they were lead to believe that they were hearing.

## The Present Study

It is this conceptual cuing that motivates the present study. While L3 learners likely detect phonetic cross-linguistic similarities in L3 input, the present study examines the role of conceptual cueing on L3 categorization at the first exposure to L3 input. Following the previous literature, which demonstrates that Spanish/English bilinguals categorize the same continuum of sounds differently depending upon the language they believe they are hearing, the present study both replicates these findings in the /p/-/b/ VOT continuum, extends them to a new F2 /i/-/u/ continuum, and examines whether one of these boundaries are used to categorize sounds in a new language, unknown to the speaker. Additionally, the present study is concerned with whether the third language has a well known relationship with previously known languages, and if this affects sound categorization.
  The distinction between /p/ and /b/ in Spanish is one of "true voicing", in which a positive VOT can be observed for the "voiceless" /p/, and a positive VOT is observed in the "voiced" /b/. English, on the other hand, has short and long lag distinction between /p/ and /b/ which is caused by a difference in aspiration. That is, both /p/ and /b/ may be phonetically voiceless (have a positive VOT), but nevertheless be categorized as voiced and voiceless by English speakers. French VOT varies by dialect. Canadian French, for instance, is more similar to English in that it has the short-lag/long lag distinction, whereas European French maintains its lead/short lag distinction [@caramazza_voice_1974]. Hungarian is considered to be a true-voicing language [@lisker_cross-language_1964], in which a /b/-/p/ distinction is both phonetically and phonemically voiced or voiceless. 
  In the case of the present study, the participants do not have any exposure to a third language, and the language specific phonetic criteria for both voicing and vowel distinctions are of minimal importance, since the participants will not have had exposure to L3 input. This design was chosen to inform how L3 learners will perceive input in the L3 when they engage it for the very first time, and which of their two phonemic boundaries they will utilize.
  As detailed previously, the entirety of the empirical support for the TPM comes from studies which involve two romance languages and English. For this reason, the present study will include a group of this kind (the French L3 group), in which it is expected that the participants will account for these stimuli using their Spanish boundaries, which would partially support TPM, but provide an alternative motivation, namely top-down conceptual cueing, that is behind L3 users choice of language in situations in which typological relationship is evident. Alternatively, an L3 Hungarian group will also be included, in which the typological relationship between the languages of this group (English, Spanish, and Hungarian) is unclear. In this case, the categorization of what participants believe are Hungarian sounds should vary in the prediction of the TPM. That is, some speakers should categorize the L3 sounds using their L1, and others their L2. Alternatively, it is possible that the Hungarian group will use their second language boundaries to categorize the L3 continua.

#### Research questions

RQ1:	Assuming that bilinguals display the same double phonemic boundary demonstrated in previous studies, which of the two language-specific phonemic boundaries will participants perceive when they are introduced to the same VOT/f2 continuum in a “third language?”
I hypothesize that that the perception of the “third language” will be modulated by the participants’ order of acquisition in the event that language relatedness is unclear (the Hungarian group), and, alternatively, that participants’ perception will be modulated by their knowledge about the languages relatedness when a clear connection is made between languages at an individual level (the French group). In particular, I predict that, when hearing the same continua, participants who are led to believe they are hearing French will perceive it more closely to Spanish than English (as demonstrated by their phonemic boundaries). On the other hand, I predict that participants who believe they are hearing Hungarian will categorize sounds by using the phonemic boundary of their second language (L2). These predictions are based on findings in previous research on L3 transfer which have demonstrated the default transfer of the L2 to the L3 in language combinations which do not involve romance languages. 

RQ2:	Will categorizations of what participants believe to be Hungarian or French result in different "language specific" categorization?
Previous research in L3 syntactic transfer [@baauw_cognitive_2013; @puig-mayenco_systematic_2020; @rothman_typological_2010] has suggested that language transfer to the L3 is not determined by order of acquisition, but by typological similarity of the languages involved. The present study offers an alternative account for the data in previous studies by suggesting that the results of previous studies can be explained by a conceptual priming effect that biases the L3 learner to utilize a particular system in order to parse L3 input. Based on evidence from a meta-analysis of empirical studies [@puig-mayenco_systematic_2020] in L3 transfer, it appears that there is no evidence of typological transfer outside of Romance language combinations when the third language is not a Romance language. Based on this evidence, the present study predicts that Spanish/English bilinguals who believe that they are hearing a romance L3 will be more likely to show bias to perceive it as Spanish-like, rather than English like. The design of the present study allows for the teasing apart of typological similarity 

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
