# Results 

## Descriptive statstics 

Tables 1 and 2 provide descriptive statistics of the crossover boundaries per group per language. The crossover boundary is the continuum step at which the probability of choosing both options (/p/ or /b/, or /i/ or /u/) was 50% and was calculated by fitting a logistic regression model to each participant in each language mode session. The crossover boundaries were standardized to z-scores in order to allow for comparability accross all language modes and over features (between stops and vowels). 

```{r}
apa_table(
  desc_vowels
  , caption = "Descriptive statistics of vowel crossovers by group"
  , escape = TRUE
)
```


```{r}
apa_table(
  desc_stops
  , caption = "Descriptive statistics of stops crossovers by group"
  , escape = TRUE
)
```

## Question 0: Was the double phonemic boundary effect replicted in stops and extended to vowels? 

In order to answer the research questions of the present study, it was vital to replicate previous findings of double phonemic boundary effect in Spanish-English bilinguals. Thus, an implicit research question in this study was whether Spanish-English bilinguals would show evidence of a double phonemic boundary in stops, as in previous studies, and, newly, whether this effect could be extended to vowels. Of course, without finding two language specific language mode-driven categorizations of English and Spanish, conclusions cannot be drawn as to whether bilinguals approach a third language using their first or second language perceptual routines. The double phonemic boundaries were assessed per language using a series of paired t-tests. The results showed that the English L1 group showed a double phonemic boundary effect in stops (`r el1ses$statistic`), but not in vowels (`r el1ves$statistic`). The Spanish L1 group showed evidence of the double phonemic boundary in both stops (`r sl1ses$statistic`) and vowels (`r sl1ves$statistic`). Figure 2 shows a graphical representations of the categorizations of vowels (figure 2) and stops (figure 3) in both languages in both L1 groups. 

```{r}

all_data %>% 
  filter(exp == "vowels", !(participant %in% vowels_remove)) %>% 
  filter(language == "english"| language == "spanish") %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 2: Spanish/English vowel boundaries per L1 group") 

```

```{r}

all_data %>% 
  filter(exp == "stops", !(participant %in% stops_remove)) %>% 
  filter(language == "english"| language == "spanish") %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 3: Spanish/English Stop boundaries per L1 group") 

```

## RQ1: which language-specific phonemic boundaries will participants perceive when they are introduced to the same VOT/F2 continuum in a “third language?” 

Due to the results of the initial series of paired t-tests, only the results of the stops in both L1 groups and the vowels in the Spanish groups were useful in answering this question, since English vowels violated the necessary assumption that all groups would show double phonemic boundary effects. A series of three paired t-tests are reported per each of the four participant groups. In each group, the paired t-tests serve to provide evidence that language categorizations are independent samples (if the p-value comparing two languages is significant) or not independent samples (if the p-value comparing two languages is not significant). The three t-tests in each group test the independence of their L1 and L2 categorization (in all 4 groups), the categorization of their L1 and their L3 (English or Spanish with Hungarian or French), and their L2 and their L3 (also English or Spanish with Hungarian or French). By having all three of the t-tests together, it can more easily be argued that one specific language has greater influence on L3 categorization if, first the Spanish and English paired t-test is significant, and one of the two remaining t-tests is significant and the other is not (i.e. if the L2 to L3 t-test is significant but the L1 to L3 t-test is not, that is taken as evidence that the L2 and the L3 are independent samples, but that the L1 and the L3 are not).

### Stops English L1 French L3 group  
English L1 vowels did not show a group wide L1-L2 double phonemic boundary effect, and subsequent analyses in the English L1 groups focuses on stops.Despite the group-wide differences in the English L1 group, double phonemic boundary effects were not found in the individual English L1 groups. In stops, paired t-tests found non-significant results, interpreted as similar categorizations in English-Spanish categorization (`r el1fess$statistic`), Spanish-French categorization (L2 to L3) (`r el1ssf$statistic`) and English-French categorization (L1 to L3) (`r el1sef$statistic`). 

### Stops English L1 Hungarian L3 group

The English-Hungarian stops were found to be distinct. English-Hungarian categorization (L1 to L3) (`r el1seh$statistic`), but not the English-Spanish categorization (`r el1hess$statistic`) nor Spanish-Hungarian categorization (L2 to L3) (`r el1ssh$statistic`). 

### Stops Spanish L1 French L3 group

No categorization differences were found in Spanish L1 speakers assigned to the "French" group in stops. There was no significant English-Spanish categorization difference (`r sl1fess$statistic`), no significant English-French categorizations difference (`r sl1sef$statistic`) and no significant Spanish-French categorization difference (`r sl1ssf$statistic`).

### Vowels Spanish L1 French L3 group

Evidence of a double phonemic Spanish-English boundary was not found in this group (`r sl1fes$statistic`). Likewise, Spanish-French categorizations had no evidence of being distinct (`r sl1vsf$statistic`). There was, however, a significant distinction in English-French categorizations (`r sl1vef$statistic`). 


### Stops Spanish L1 Hungarian L3 group

In this group, the English-Spanish categorization difference was significant (`r sl1hess$statistic`), implying that this group showed evidence of the double phonemic boundary. The L2 and L3 English-Hungarian categorizations were also found to be categorized differently (`r sl1seh$statistic`), whereas the L1 and the L3 Spanish-Hungarian categorizations (`r sl1ssh$statistic`) were not categorized differently. 

### Vowels Spanish L1 Hungarian L3 group

The same trend is present in vowels as in stops for the Spanish L1 Hungarian L3 group. A significant English-Spanish categorization t-test (`r sl1hes$statistic`) reveals a double phonemic boundary effect. There is a significant difference in L2-L3 English-Hungarian categorizations (`r sl1veh$statistic`) ,and there is not significant L1-L3 Spanish-Hungarian categorization distinction (`r sl1vsh$statistic`).


## RQ2: Did the L3 groups categorize the continua differently?

T-tests of independent samples revealed that L3 groups did not categorize their respective continua differently if they were assigned to the French group or the Hungarian group. The L3 French group with Spanish L1 and the L3 Hungarian group with a Spanish L1 did not categorize their respective L3s differently in vowels (`r sl1vfh$statistic`) or stops (`r sl1sfh$statistic`). The same evidence was found the the English L1 groups; they categorized L3 sounds similarly, whether they were in the L3 French group or Hungarian group in both vowels (`r el1vfh$statistic`) and stops (`r el1sfh$statistic`). 
 

## Post-hoc analysis 

Based on the findings that the L3 continua were not categorized differently by the same L1 group, a post-hoc series of paired t-tests was carried out in which the L3 groups were pooled together. This was done with the aim of increasing sample size and to show that L1 boundaries and L3 boundaries in the Spanish L1 group are not categorized differently (interpreted as a non-significant p-value), and that the L2 and the L3 and the L1 and the L2 are categorized differently (both should have a significant p-value). This trend was also found in English L1 group stop categorizations.

### Spanish L1 group pooled stops 

These results suggest that L1 and L3 are categorized similarly in the stops of the Spanish L1 group, (`r pooled_stops_sl1_sl3$statistic`), together with evidence that the L3s are not significantly differentiated between groups, and that this group shows evidence a double phonemic boundary (`r pooled_stops_sl1_se$statistic `). On the other hand, L2 English is categorized differently than the L3 (`r pooled_stops_sl1_el3$statistic`). Figure 3 is a graphical representation of the crossover boundaries of stops in the pooled data of both Spanish L1 groups.

```{r}

all_data %>% 
  filter(exp == "stops", !(participant %in% stops_remove)) %>% 
  filter(l1 == "Spanish") %>% 
  mutate(language = ifelse(as.character(language) == "french", "l3", as.character(language))) %>% 
  mutate(language = ifelse(as.character(language) == "hungarian", "l3", as.character(language))) %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /p/") + xlab("VOT") + 
  ggtitle("Figure 3: Spanish L1 group pooled stops") 


```


### Spanish L1 group pooled vowels 

The results of the vowels mirror those found in stops in the Spanish L1 group. First, there was evidence of a double phonemic boundary (`r pooled__vowel_sl1_se$statistic`). The L1 and the L3 were categorized similarly (`r pooled__vowel_sl1_sl3$statistic`), and the L2 and the L3 were categorized distinctly (`r pooled__vowel_sl1_el3$statistic`). Figure 4 is a graphical representation of the crossover boundaries of vowels in the pooled data of both Spanish L1 groups.


```{r}

all_data %>% 
  filter(exp == "vowels", !(participant %in% vowels_remove)) %>% 
  filter(l1 == "Spanish") %>% 
  mutate(language = ifelse(as.character(language) == "french", "l3", as.character(language))) %>% 
  mutate(language = ifelse(as.character(language) == "hungarian", "l3", as.character(language))) %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 3: Spanish L1 group pooled vowels") 

```

### English L1 group pooled stops 

In the case of the English L1 group, only the stops showed evidence of a double phonemic boundary modulated by language mode. This Spanish-English double phonemic boundary effect does not quite persist in those who completed all phases of the experiment (`r pooled_stops_el1_se$statistic`). The English L1 group patterned oppositely in comparison to the Spanish L1 group, since there was a difference between their L1 (English) and L3 categorizations (`r pooled_stops_el1_el3$statistic`). On the other hand, L2 categorizations and L3 categorizations were not significantly different in this group (`r pooled_stops_el1_sl3$statistic`). These finding suggests a greater degree of influence from the L2 on L3 stop categorizations in the L1 group. 


```{r}

all_data %>% 
  filter(exp == "stops", !(participant %in% stops_remove)) %>% 
  filter(l1 == "English") %>% 
  mutate(language = ifelse(as.character(language) == "french", "l3", as.character(language))) %>% 
  mutate(language = ifelse(as.character(language) == "hungarian", "l3", as.character(language))) %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /p/") + xlab("VOT") + 
  ggtitle("Figure 5: English L1 group pooled stops") 

```



