---
title: "Preliminary Results"
output: github_document
---


```{r, echo=FALSE, include=FALSE}

library("here")
library("dplyr")
library("tidyr")
library("stringr")
library("readr")
library("ggplot2")
library("purrr")
library("fs")
library("patchwork")

grand_n = read.csv(here("data", "tidy", "grand_n.csv"))

total_participants = sum(grand_n$n)

```

```{r, echo=FALSE, include=FALSE}



tidy_2afc <- . %>% 
  map_dfr(read_csv, .id = "source", col_types = cols(.default = "c")) %>% 
  select(step_cont = stimuli, participant, date, expName, resp_2.keys, resp_3.keys) %>% 
  mutate(exp = if_else(is.na(resp_2.keys), "stops", "vowels"), 
         step_cont = str_remove(step_cont, "stim/VC_"), 
         step_cont = str_remove(step_cont, "stim2/pa"), 
         step_cont = str_remove(step_cont, "\\+"), 
         step_cont = str_remove(step_cont, ".wav"), 
         step_cont = as.numeric(step_cont), 
         response = case_when(
           resp_2.keys == "left"  ~ 0, 
           resp_2.keys == "right" ~ 1, 
           resp_3.keys == "left"  ~ 1, 
           resp_3.keys == "right" ~ 0)) %>% 
  filter(nchar(as.character(participant)) == 24)

# -----------------------------------------------------------------------------




# Tidy English data -----------------------------------------------------------

english <- dir_ls(here("raw_data", "english_2afc"), regexp = "\\.csv$") %>%
  tidy_2afc() %>% 
  mutate(language = "english")

english_vowels <- english %>% 
  filter(!is.na(resp_2.keys)) %>% 
  select(participant, date, expName, resp_2.keys, exp, language, 
         step_cont, response)

english_stops <- english %>% 
  filter(!is.na(resp_3.keys)) %>% 
  select(participant, date, expName, resp_2.keys, exp, language, 
         step_cont, response)

# -----------------------------------------------------------------------------



# Tidy Spanish Data -----------------------------------------------------------

spanish <- dir_ls(here("raw_data", "spanish_2afc"), regexp = "\\.csv$") %>% 
  tidy_2afc() %>% 
  mutate(language = "spanish")

spanish_vowels <- spanish %>% 
  filter(!is.na(resp_2.keys)) %>% 
  select(participant, date, expName, resp_2.keys, exp, language, 
         step_cont, response)

spanish_stops <- spanish %>% 
  filter(!is.na(resp_3.keys)) %>% 
  select(participant, date, expName, resp_3.keys, exp, language, 
         step_cont, response)

# -----------------------------------------------------------------------------




## Tidy French data -----------------------------------------------------------

french <- dir_ls(here("raw_data", "french_2afc"), regexp = "\\.csv$") %>%
  tidy_2afc() %>% 
  mutate(language = "french")

french_vowels <- french %>% 
  filter(!is.na(resp_2.keys)) %>% 
  select(participant, date, expName, resp_2.keys, exp, language, 
         step_cont, response)

french_stops <- french %>% 
  filter(!is.na(resp_3.keys)) %>% 
  select(participant, date, expName, resp_3.keys, exp, language, 
         step_cont, response)

# -----------------------------------------------------------------------------




# Tidy Hungarian data ---------------------------------------------------------

hungarian <- dir_ls(here("raw_data", "hungarian_2afc"), regexp = "\\.csv$") %>% 
  tidy_2afc() %>% 
  mutate(language = "hungarian") 


hungarian_vowels <- hungarian %>% 
  filter(!is.na(resp_2.keys)) %>% 
  select(participant, date, expName, resp_2.keys, exp, language, 
         step_cont, response)

hungarian_stops <- hungarian %>% 
  filter(!is.na(resp_3.keys)) %>% 
  select(participant, date, expName, resp_3.keys, exp, language, 
         step_cont, response)

# -----------------------------------------------------------------------------




# Combine data ----------------------------------------------------------------

# Add missing participant labels
lhq1 <- read_csv("./raw_data/lhq/LHQ 3.0 raw result (1).csv") %>% 
  janitor::clean_names() %>% 
  select(l1 = 9, l2 = 15, participant = 1, prolific = 2)
lhq2 <- read_csv("./raw_data/lhq/LHQ 3.0 raw result (2).csv") %>% 
  janitor::clean_names() %>% 
  select(l1 = 9, l2 = 15, participant = 1, prolific = 2)
lhq3 <- read_csv("./raw_data/lhq/LHQ 3.0 raw result (3).csv") %>% 
  janitor::clean_names() %>% 
  select(l1 = 9, l2 = 15, participant = 1, prolific = 2)
lhq5 <- read_csv("./raw_data/lhq/LHQ 3.0 raw result (5).csv") %>% 
  janitor::clean_names() %>% 
  select(l1 = 9, l2 = 15, participant = 1, prolific = 2)
lhq6 <- read_csv("./raw_data/lhq/LHQ 3.0 raw result (6).csv") %>% 
  janitor::clean_names() %>% 
  select(l1 = 9, l2 = 15, participant = 1, prolific = 2)

lhq <- bind_rows(lhq1, lhq2, lhq3, lhq5,lhq6)

missing_ids<- read_csv("./raw_data/lhq/prolifictolhq.csv")

lhq_sub = lhq %>% 
  filter(nchar(as.character(participant)) == 5)

lhq_main = lhq %>% 
  filter(nchar(as.character(participant)) == 24)

missing_ids = missing_ids %>% 
  janitor::clean_names() %>% 
  select(participant = 2, prolific_id = 1) %>% 
  left_join(., lhq_sub, by = "participant") %>% 
  select(l1, l2, participant, prolific_id) %>% 
  select(l1 = 1, l2 = 2, participant = 3 ,prolific = 4)


lhq = lhq_main %>% 
  rbind(., missing_ids) %>%
  filter(nchar(as.character(prolific)) == 24) %>% 
  select(l1 = 1, l2 = 2, prolific = 3, participant = 4)
  

all_stops  <- bind_rows(
  english_stops, hungarian_stops, spanish_stops, french_stops) %>% 
  left_join(., lhq, by = "participant") %>% 
  mutate(group = if_else(l2 == "Spanish", "ES", "SE")) %>% 
  filter(!is.na(group)) %>% 
  write_csv(here("data", "tidy", "2afc_stops_tidy.csv"))

all_vowels <- bind_rows(
  english_vowels, spanish_vowels, french_vowels, hungarian_vowels) %>% 
  left_join(., lhq, by = "participant") %>% 
  mutate(group = if_else(l2 == "Spanish", "ES", "SE")) %>% 
  filter(!is.na(group)) %>% 
  write_csv(here("data", "tidy", "2afc_vowels_tidy.csv"))

assign_group_vowels = all_vowels %>%
  filter(language == "french" | language == "hungarian") %>%
  mutate(l3group = if_else(language == "french", "f", "h")) %>% 
  select(participant, l3group)


assign_group_stops = all_stops %>%
  filter(language == "french" | language == "hungarian") %>%
  mutate(l3group = if_else(language == "french", "f", "h")) %>% 
  select(participant, l3group)

all_vowels = all_vowels %>% 
  
  left_join(., assign_group_vowels, by = "participant") 

all_stops = all_stops %>% 
  left_join(., assign_group_stops, by = "participant")
ebv = all_vowels %>% 
  filter(., language == "english") %>% 
  group_by(participant) %>%
  summarise(ebv = mean(response))

sbv = all_vowels %>% 
  filter(., language == "spanish") %>% 
  group_by(participant) %>%
  summarise(sbv = mean(response))

fbv = all_vowels %>% 
  filter(., language == "french") %>% 
  group_by(participant) %>%
  summarise(l3bv = mean(response))

fbv = fbv %>%
  filter(!is.na(l3bv)) %>% 
  mutate(l3group = "f")

hbv = all_vowels %>% 
  filter(., language == "hungarian") %>% 
  group_by(participant) %>%
  summarise(l3bv = mean(response))

hbv = hbv %>%
  filter(!is.na(l3bv)) %>% 
  mutate(l3group = "h")


l3bv = rbind(fbv,hbv)

##  Merge LHQ data with 2afc data, filter out those who did not do all tasks. 

analyze_vowels = left_join(ebv, sbv, on = "participant") %>% 
  left_join(., l3bv, on = "participant") %>% 
  left_join(lhq, on = "participant") %>% 
  filter(!is.na(sbv)) %>% 
  filter(!is.na(l3bv)) %>% 
  mutate(group = if_else(l2 == "Spanish", "ES", "SE"))




## Descriptive stats of mean boundaries per group for vowels

analyze_vowels %>% 
  group_by(group) %>%
  summarise(mean_ebv = mean(ebv), sd_ebv = sd(ebv), mean_sbv = mean(sbv), sd(sbv), n = n())


### Filter for SE group and exclude participants who chose the same answer 90 percent of the time.

se_descriptive = analyze_vowels %>% 
  filter(group == "SE") %>% 
  mutate(es_effect = ebv - sbv) %>%
  filter(ebv < .9)

seh_descriptive = analyze_vowels %>% 
  filter(group == "SE") %>% 
  filter(l3group == "h") %>%
  summarize(mean_ebv = mean(ebv), mean_sbv = mean(sbv), mean_l3bv = mean(l3bv), n = n()) %>% 
  mutate(group = "seh")

sef_descriptive = analyze_vowels %>% 
  filter(group == "SE") %>% 
  filter(l3group == "f") %>% 
  summarize(mean_ebv = mean(ebv), mean_sbv = mean(sbv), mean_l3bv = mean(l3bv), n = n()) %>% 
  mutate(group = "sef")

### Filter for ES group and exclude participants who chose the same answer 90 percent of the time. 

es_descriptive = analyze_vowels %>% 
  filter(group == "ES") %>% 
  mutate(es_effect = ebv - sbv) %>%
  filter(ebv < .9)

esh_descriptive = analyze_vowels %>% 
  filter(group == "ES") %>% 
  filter(l3group == "h") %>% 
  summarize(mean_ebv = mean(ebv), mean_sbv = mean(sbv), mean_l3bv = mean(l3bv), n = n()) %>% 
  mutate(group = "esh")

esf_descriptive = analyze_vowels %>% 
  filter(group == "ES") %>% 
  filter(l3group == "f") %>% 
  summarize(mean_ebv = mean(ebv), mean_sbv = mean(sbv), mean_l3bv = mean(l3bv), n = n()) %>% 
  mutate(group = "esf")

ebs = all_stops %>% 
  filter(., language == "english") %>% 
  group_by(participant) %>%
  summarise(ebs = mean(response))

sbs = all_stops %>% 
  filter(., language == "spanish") %>% 
  group_by(participant) %>%
  summarise(sbs = mean(response))

fbs = all_stops %>% 
  filter(., language == "french") %>% 
  group_by(participant) %>%
  summarise(l3bs = mean(response))

fbs = fbs %>%
  filter(!is.na(l3bs)) %>% 
  mutate(l3group = "f")

hbs = all_stops %>% 
  filter(., language == "hungarian") %>% 
  group_by(participant) %>%
  summarise(l3bs = mean(response))

hbs = hbs %>%
  filter(!is.na(l3bs)) %>% 
  mutate(l3group = "h")


l3bs = rbind(fbs,hbs)

##  Merge LHQ data with 2afc data, filter out those who did not do all tasks. 

analyze_stops = left_join(ebs, sbs, on = "participant") %>% 
  left_join(., l3bs, on = "participant") %>% 
  left_join(lhq, on = "participant") %>% 
  filter(!is.na(sbs)) %>% 
  filter(!is.na(l3bs)) %>% 
  mutate(group = if_else(l2 == "Spanish", "ES", "SE"))




## Descriptive stats of mean boundaries per group for stops

analyze_stops %>% 
  group_by(group) %>%
  summarise(mean_ebs = mean(ebs), sd_ebs = sd(ebs), mean_sbs = mean(sbs), sd(sbs), n = n())


### Filter for SE group and exclude participants who chose the same answer 90 percent of the time.

se_descriptive_s = analyze_stops %>% 
  filter(group == "SE") %>% 
  mutate(es_effect = ebs - sbs) %>%
  filter(ebs < .9)

seh_descriptive_s = analyze_stops %>% 
  filter(group == "SE") %>% 
  filter(l3group == "h") %>% 
  summarize(mean_ebv = mean(ebs), mean_sbs = mean(sbs), mean_l3bs = mean(l3bs), n = n())

sef_descriptive_s = analyze_stops %>% 
  filter(group == "SE") %>% 
  filter(l3group == "f") %>% 
  summarize(mean_ebs = mean(ebs), mean_sbs = mean(sbs), mean_l3bs = mean(l3bs), n = n())

### Filter for ES group and exclude participants who chose the same answer 90 percent of the time. 

es_descriptive_s = analyze_stops %>% 
  filter(group == "ES") %>% 
  mutate(es_effect = ebs - sbs) %>%
  filter(ebs < .9)

esh_descriptive_s = analyze_stops %>% 
  filter(group == "ES") %>% 
  filter(l3group == "h") %>% 
  summarize(mean_ebs = mean(ebs), mean_sbs = mean(sbs), mean_l3bs = mean(l3bs), n = n())

esf_descriptive_s = analyze_stops %>% 
  filter(group == "ES") %>% 
  filter(l3group == "f") %>% 
  summarize(mean_ebs = mean(ebs), mean_sbs = mean(sbs), mean_l3bs = mean(l3bs), n = n())



```
## Overview

Here I will briefly show the results of the 2afc task by 4 groups, including both orders of acquisition (Spanish-English and English-Spanish) who were told that they were hearing either French or Hungarian (the variable L3 group).

**Note** for the boundaries, I calculated the mean responses (from 0-1), rather than using the cross-over function, due to lack of ability. I also used this as exclusion criteria; if a participant has a response average over .9 they were omitted, since this means they probably hit the same button most of the time and did not pay attention. 

A total of `r total_participants` completed all portions of the experiment, and met the eligibility criteria for the study. There were a total of `r esf_descriptive %>% pull(n)` English-Spanish bilinguals assigned to the French group, and `r esh_descriptive %>% pull(n)` assigned to the Hungarian group. `r sef_descriptive %>% pull(n)` Spanish-English bilinguals were assigned to the French group and `r seh_descriptive %>% pull(n)` were assigned to the Hungarian group. 

## L1-L2 Vowels: the double phonemic boundary

The results of the present study found a double phonemic boundary between the vowels /i/ and /u/ in the Spanish-English group, but not the English-Spanish group. In order to establish which perceptual boundary the participants use when they encounter speech in a new language, it is first important to replicate previous studies that have demonstrated double-phonemic boundary effects in bilinguals. Of course, the presence of two boundaries is a prerequisite to answer the questions posed by the present study. As a result, the results of the ES group are essentially null, in that they did not show a double phonemic boundary, and it cannot be determined which of their two languages they employ when they encounter L3 speech. On the other hand, evidence was found for a double phonemic boundary the SE group, which allows for the comparison of this group's boundaries to their L3 categorizations. In order to determine whether both groups show evidence of a double phonemic boundary, paired T-tests were carried out. 

## T-tests for L1-L2 vowels

**Paired T-test for the ES group's L1-L2 vowel categorization**
```{r, echo=FALSE}

t.test(es_descriptive$ebv, es_descriptive$sbv, paired = TRUE)

```
**Paired T-test for the SE group's L1-L2 vowel categorization**
```{r, echo=FALSE}

t.test(se_descriptive$ebv, se_descriptive$sbv, paired = TRUE)

```

## T-tests for L1-L2 stops

**Paired T-test for the ES group's L1-L2 stop categorization**
```{r, echo=FALSE}

t.test(es_descriptive_s$ebs, es_descriptive_s$sbs, paired = TRUE)

```
**Paired T-test for the SE group's L1-L2 stop categorization**
```{r, echo=FALSE}

t.test(se_descriptive_s$ebs, se_descriptive_s$sbs, paired = TRUE)

```

## Plots 

Plots show the s-shaped curves of the responses by group. The x-axis shows the step in the continuum and the y axis shows the percentage response either /p/ for stops or /u/ for vowels. 


#### Plot L3 boundaries by order of acquisition 

```{r, echo=FALSE}

# Plot L3 boundaries by order of acquisition 
ggplot(data = analyze_vowels, aes(x = group, y = l3bv, color = l3group)) + 
  geom_boxplot() + ylab("L3 boundary")

```

#### Plot boundary per language per group 

```{r, echo=FALSE}

analyze_vowels %>% 
  pivot_longer(c(`ebv`, `sbv`, `l3bv`), names_to = "language", values_to = "step") %>% 
  ggplot(., aes(x = group, y = step, color = language)) + geom_boxplot()
```

### Plot S-shaped curves per group 

#### SEH vowels
```{r, echo=FALSE}

all_vowels %>%
  filter(group == "SE") %>% 
  filter(l3group == "h") %>%
  ggplot(., aes(x = step_cont, y = response, color = language)) + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  labs(title = "L1 Spanish L2 English Hungarian Vowels")
```

#### SEF vowels
```{r, echo=FALSE}

all_vowels %>%
  filter(group == "SE") %>% 
  filter(l3group == "f") %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  labs(title = "L1 Spanish L2 English French Vowels")

ttest_seh = analyze_vowels %>% 
  filter(l3group == "f") %>% 
  filter(group == "SE") 
  
```

#### Facet SE group by l3 group

```{r, echo=FALSE}
all_vowels %>%
  filter(group == "SE") %>% 
  filter(!is.na(l3group)) %>%
  ggplot(., aes(x = step_cont, y = response, color = language)) + 
  facet_grid(. ~ l3group) + geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  labs(title = "L1 Spanish L2 English Vowels")
```

#### ESH vowels 
```{r, echo=FALSE}

all_vowels %>%
  filter(group == "ES") %>% 
  filter(l3group == "h") %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  labs(title = "L1 English L2 Spanish Hungarian Vowels")
```

#### ESF vowels 
```{r, echo=FALSE}

all_vowels %>%
  filter(group == "ES") %>% 
  filter(l3group == "f") %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  labs(title = "L1 English L2 Spanish French Vowels")

```

#### Facet ES group by l3 group
```{r, echo=FALSE}

all_vowels %>%
  filter(group == "ES") %>% 
  filter(!is.na(l3group)) %>%
  ggplot(., aes(x = step_cont, y = response, color = language)) + 
  facet_grid(. ~ l3group) + geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  labs(title = "L1 English L2 Spanish Vowels")


```


