---
title: "The categorization of L3 sounds in Spanish/English bilinguals during the initial stage of L3 acquisition"
subtitle: 'Qualifying Paper Defense Spring 2021'
author  : "Kyle Parrish"
date    : "Rutgers University</br>Spring 2021</br>Last update: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: assets
    css: ["rutgers"]
    nature:
      beforeInit: 
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE, fig.height = 6, fig.width = 7)
```

```{r, echo=FALSE, include=FALSE}
source(here::here("scripts", "07_manuscript_data.R"))
```

# Overview

- Broad questions

- Literature review

- Research questions

- Methods

- Results 

- Discussion 

---

# Broad questions 

- When a bilingual learns third language, how does it work?

--

 - Which language known by a bilingual impacts L3 acquisition? 
 
--

 - Does it matter which languages a bilingual knows, and which third language they are learning?

--

 - Should we expect different linguistic domains, such as morphosyntax and phonology, to be influenced in the same way by previously learned language in an L3 learner?

---

# Literature Review

**Models of L3A**

.pull-left[
- There has been a focus on morphosyntax in L3 models:

- The **Typological Primacy Model** (Rothman 2010, 2011, 2013, 2015) suggests that perceived typological similariy causes either the L1 or the L2 to transfer to the L3, but not both. 

- The **L2 Status Factor** (Bardel & Falk, 2007) predicts that the second learned language will influence the L3 by default, and "block" access to the L1. 

- The **Linguistic Proximity Model** (Westergaard et al. 2017) posits that structure-by-structure influence is possible, but does not predict when it occurs. 
]

--

.pull-right[
- I test these models' predictions in L3 phonology

 - Specifically, it was of interest how L3 learners would perceive sounds in their third language they very first time they were exposed to it. 
]

---

# Literature Review

**Bilingual Phonology**

- To examine this question, I adapted methods used in studies in bilingual phonology (Lozano-Argüelles et al. 2020; Casillas & Simonet, 2018; Gonzales et al., 2019).
 
 - These studies found that bilinguals show evidence of two, language-specific categorization routines.

--

 - The studies used a voicing continuum, realized as the beginning of a pseudoword (pafri-bafri), and presented participants with 2 alternative forced choice. 

 - Vitally, participants were told that they would be listening to rare words in either English or Spanish, but categorized the same continuum in each session. 

--

- The results showed that the categorization of the same continuum, by the same participants, differed according to the language they believed they were hearing. 

- This finding has been referred to as the **double-phonemic boundary effect**, in which two language specific categorizations of similar cross-linguistic phonemes are possible in the same bilingual speaker. 

---

# Research questions

- Building on previous DPBE studies, I added a third continuum in a "third language", which I told participants was either French or Hungarian. 
 - the stop continuum used in previous studies was used here.
 - A vowel continuum was also created and used in order to evaluate the predictions of full transfer models. 
 
--

**RQ1:** Will the double phonemic boundary effect found in previous studies be replicated and extended to vowels?

--

**RQ2:** Will the participants' categorizations of what they believe are the L3 continua resemble their categorizations of either their L1 or their L2? 

--

**RQ3:** Will categorizations of what participants believe to be Hungarian or French result in different “language specific” categorization between L3 groups?

--

It was predicted that participants would categorize "L3 French" as similar to Spanish, but "L3 Hungarian" as similar to their L2. 

---

# Methods

**Participants**

- All participants were recruited from the online platform Prolific and completed all parts of the experiment online. 

- 100 participants took part in at least one part of the study.
 - The data of 16 participants was removed for not completing all blocks of the study.

- Two groups of Spanish-English bilinguals were created: English L1 (n = 36) and Spanish L1 groups (n = 55).

--

- These groups were further randomly subdivided into L3 groups: either the French or Hungarian.
 - Spanish L1 Hungarian (n = 31)
 - Spanish L1 French (n = 18)
 - English L1 Hungarian (n = 17)
 - English L1 French (n = 17)

- Thus, each participant categorized the same continua in three distinct sessions.

---

# Methods

**Materials**

- The participants completed three 2afc tasks 
 - Programmed in Psychopy (Pierce et al., 2019)

- Language History Questionnaire (LHQ 3.0, Li et al., 2020) 
 - Gathered language L1 group and language background

- lexTALE proficiency test in their L2 (Izur et al., 2016; Lemhöfer & Broersma, 2016))
 - Provided a continuous measure of proficiency 
 
---

# Methods

**Procedure**

- The participants were told they would be listening to the beginnings of rare words in English, Spanish, and a third language they do not know. 

- The participants heard one word at a time, which was randomly drawn from the continua.
 - After each word, they would choose between either pafri-bafri or ifri-ufri by pressing the left or right arrow key.
 - The participants first categorized the vowel continuum, followed by the voicing continuum.

- Vitally, the participants did not complete all 3 language sessions in a row.

- The LHQ and lexTALE were given prior to the beginning of the 2afc sessions. 

---

# Results 

**Analysis**

- The responses of each continuum were then analyzed by fitting logistic regression models to each participant in each continuum.
 - This method created a cross-over boundary for each participant in each language, in which the probability of choosing either of the given options was 50%.

- Using these crossover boundaries, a series of paired t-tests were carried out in order to determine whether the distribution of Spanish and English categorizations were distinct in the same participants.

- Based on the results, post-hoc paired tests of equivalence were also carried out.

---

# Results

**Spanish-English categorizations**

- Spanish and English were categorized differently in 3 out of 4 possible combinations:

 - The English L1 group showed a double phonemic boundary effect in stops (t(24)=2.24, p=.035), but not in vowels (t(35)=-0.72, p=.477). 

 - The Spanish L1 group showed evidence of the double phonemic boundary in both stops (t(51)=3.13, p=.003) and vowels (t(49)=3.77, p<.001)

---

# Results

.pull-left[
**Spanish-English categorizations**

- Figures 2 and 3 show a graphical representations of the categorizations of vowels (figure 2) and stops (figure 3) in both languages in both L1 groups.

- A small effect can be seen in the Spanish L1 group (right graph)
]

.pull-right[
```{r}
all_data %>% 
  filter(exp == "vowels", !(participant %in% vowels_remove)) %>% 
  filter(language == "english"| language == "spanish") %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 2: Spanish/English vowel boundaries per L1 group") 
```
]

---

# Results

.pull-left[
**Spanish-English categorizations**

- Figure 3 shows the categorizations of stops in both L1 groups. 

- Both groups showed evidence of a Spanish-English double phonemic boundary in stops. 
]

.pull-right[
```{r}
all_data %>% 
  filter(exp == "stops", !(participant %in% stops_remove)) %>% 
  filter(language == "english"| language == "spanish") %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 3: Spanish/English Stop boundaries per L1 group") 
```
]

---

# Results

**L3 categorizations**

- T-tests of the crossovers revealed that L3 groups did not categorize the continua differently whether they were assigned to the Hungarian group or the French group, and regardless of their first language. 

--

- English L1: vowels (t(31.25)=0.10, p=.924) or stops (t(37.81)=0.55, p=.585).  

- Spanish L1: vowels (t(29.90)=0.57, p=.573) and stops (t(10.89)=0.28, p=.788).

---

# Results

**L3 categorizations**

- Since the L3 groups may not have categorized the L3 differently, their results were pooled in order to carry out a post-hoc analysis.

--

- The purpose of this analysis was to provide more robust evidence of the categorizations of the continua in both L1 groups. 

---

# Post-hoc analysis 

- Tests of equivalence were carried out using the `toster` function in R.

- Cohen's D was set to .5, a medium effect size, for the upper and lower equivalence bounds.

---

# Post-hoc analysis 

**Spanish L1 pooled stops** 

.pull-left[

L1 and L3 categorizations were found to be equivalent.

On the other hand, English and the L3 were not found to be equivalent and nor were Spanish and English. 

L1 - L2 (t(45) = -0.770, p = 0.223).

L1 - L3 (t(45) = -3.094, **p = 0.00169**)

L2 - L3 (t(45) = -1.103, p = 0.138)
]

.pull-right[
```{r}
all_tost_spanl1_stops %>% 
  ggplot(aes(x = diff, y = combination)) + geom_point() + 
  geom_linerange(aes(xmin = LL_CI_TOST, xmax = UL_CI_TOST)) +
  geom_vline(xintercept = all_tost_spanl1_stops$low_eqbound_dz, linetype = "dashed") + 
  geom_vline(xintercept = all_tost_spanl1_stops$high_eqbound_dz, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = .5) +
  ggtitle("Figure 6: Spanish L1 stops TOST")
```
]

---

# Post-hoc analysis 

**Spanish L1 pooled vowels**  

.pull-left[
The same trend was observed in vowels.

L1 and L3 categorizations were equivalent, other combinations were not:

L1 - L2 (t(45) = 0.536, p = 0.703)

L1 - L3 (t(45) = -2.949, **p = 0.00252**)

L2 - L3 (t(45) = 0.198, p = 0.578)
]
.pull-right[
```{r}
all_tost_spanl1_vowels %>% 
  ggplot(aes(x = diff, y = combination)) + geom_point() + 
  geom_linerange(aes(xmin = LL_CI_TOST, xmax = UL_CI_TOST)) + 
  geom_vline(xintercept = all_tost_spanl1_vowels$low_eqbound_dz, linetype = "dashed") + 
  geom_vline(xintercept = all_tost_spanl1_vowels$high_eqbound_dz, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = .5) +
  ggtitle("Figure 8: Spanish L1 Vowels TOST") + xlim(-.05,.05)
```
]

---

# Post-hoc analysis 

**English L1 pooled stops**

.pull-left[
These results were less clear. 

L2 and L3 categorizations could be equivalent, but there is not enough data to make this determination. 

L1 - L2 (t(22) = 0.427, p = 0.663)

L1 - L3 (t(22) = -0.527, p = 0.302)

L2 - L3 (t(22) = 1.441, p = 0.0818)

This finding suggests that the English L1 group might show L2 status effects and would trend oppositely of the Spanish L1 group.
]

.pull-right[
```{r}
all_tost_engl1_stops %>% 
  ggplot(aes(x = diff, y = combination)) + geom_point() + 
  geom_linerange(aes(xmin = LL_CI_TOST, xmax = UL_CI_TOST)) +
  geom_vline(xintercept = all_tost_engl1_stops$low_eqbound_dz, linetype = "dashed") + 
  geom_vline(xintercept = all_tost_engl1_stops$high_eqbound_dz, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = .5) +
  ggtitle("Figure 10: English L1 stops")
```
]

---

# Post-hoc analysis 

**L3 categorizations**

This plot shows the Spanish L1 groups categorization of vowels in the L1, the L2 and the L3. 

```{r}
all_data %>% 
  filter(exp == "vowels", !(participant %in% vowels_remove)) %>% 
  filter(l1 == "Spanish") %>% 
  mutate(language = ifelse(as.character(language) == "french", "l3", as.character(language))) %>% 
  mutate(language = ifelse(as.character(language) == "hungarian", "l3", as.character(language))) %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 4: Spanish L1 group pooled vowels") 
```

---

# Post-hoc analysis 

**L3 categorizations**

This plot shows the Spanish L1 groups categorization of stops in the L1, the L2 and the L3. 

```{r}
all_data %>% 
  filter(exp == "stops", !(participant %in% stops_remove)) %>% 
  filter(l1 == "Spanish") %>% 
  mutate(language = ifelse(as.character(language) == "french", "l3", as.character(language))) %>% 
  mutate(language = ifelse(as.character(language) == "hungarian", "l3", as.character(language))) %>% 
  ggplot(., aes(x = step_cont, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /p/") + xlab("VOT") + 
  ggtitle("Figure 5: Spanish L1 group pooled stops") 
```

---

# Discussion 

- L3 models were not supported.
 - L1 influence was found in the Spanish L1 group only

- The DPBE was replicated in stops in both groups and extended to vowels in the Spanish L1 group.

- The English L1 group did not show evidence of the DPBE in vowels 

---

# Revisiting broad questions 

- Which language known by a bilingual impacts L3 acquisition? 
 - The Spanish L1 group provided evidence that the L1 impacts L3A - but the English L1 group might not have shown this result. 

- Does it matter which languages a bilingual knows, and which third language they are learning?
 - Does it make a difference if Spanish-English bilinguals learn French versus when they learn Hungarian? 
  - No differences were found in L3 categorizations in any group, but more data is needed. 

- Should we expect different linguistic domains, such as morphosyntax and phonology, to be influenced in the same way by previously learned language in an L3 learner? 
 - Models of L3 acquisition have focused on morphosyntax, and the relationship between L3 phonology and morphosyntax remains unclear. 
  - These results suggest that phonology and morphosyntax are distinct and should have separate models. 


---

# References 

Bardel, C., & Falk, Y. (2007). The role of the second language in third language acquisition: The case of germanic syntax. Second Language Research, 23(4), 459–484. https://doi.org/10.1177/0267658307080557

Casillas, J. V., & Simonet, M. (2018). Perceptual categorization and bilingual language modes: Assessing the double phonemic boundary in early and late bilinguals. Journal of Phonetics, 71, 51–64. https://doi.org/10.1016/j.wocn.2018.07.002

Gonzales, K., Byers-Heinlein, K., & Lotto, A. J. (2019). How bilinguals perceive speech depends on which language they think they’re hearing. Cognition, 182, 318–330.

Izura, C., Cuetos, F., & Brysbaert, M. (2016). Lexical test for advanced learners of spanish. American Psychological Association. https://doi.org/10.1037/t47086-000

Lemhöfer, K., & Broersma, M. (2016). Lexical test for advanced learners of english. American Psychological Association. https://doi.org/10.1037/t47085-000

---

# References (cont.)

Lozano-Argüelles, C., Fernández Arroyo, L., Rodríguez, N., Durand López, E. M., Garrido Pozú, J. J., Markovits, J., … Casillas, J. V. (2020). CONCEPTUALLY CUED PERCEPTUAL CATEGORIZATION IN ADULT L2 LEARNERS. Studies in Second Language Acquisition, 1–16. https://doi.org/10.1017/S0272263120000273

Peirce, J., Gray, J. R., Simpson, S., MacAskill, M., Höchenberger, R., Sogo, H., … Lindeløv, J. K. (2019). PsychoPy2: Experiments in behavior made easy. Behavior Research Methods, 51(1), 195–203. https://doi.org/10.3758/s13428-018-01193-y

R Core Team. (2020). R: A language and environment for statistical computing. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from https://www.R-project.org/

Rothman, J. (2015). Linguistic and cognitive motivations for the typological primacy model (TPM) of third language (L3) transfer: Timing of acquisition and proficiency considered. Bilingualism: Language and Cognition, 18(2), 179–190. https://doi.org/10.1017/S136672891300059X

Westergaard, M., Mitrofanova, N., Mykhaylyk, R., & Rodina, Y. (2017). Crosslinguistic influence in the acquisition of a third language: The linguistic proximity model. International Journal of Bilingualism, 21(6), 666–682. https://doi.org/10.1177/1367006916648859
