# Methods

## Participants

<!-- It looks like your are pulling a vector of values (there are NAs) --> 
A total of `r language_n_man %>% pull(en)` participants took part in at least one experiment of the study and were recruited on the online platform Prolific. A total of 16 participants did not complete all blocks of the experiment, and their data was removed from the final analysis, resulting in a total of 84 participants who completed all portions of the experiment. 
In order to tease apart L2 status effects from CLI brought about by typological similarity across languages, participants were recruited with both orders of acquisition. 
That is, one group of participants spoke English as a first language and Spanish as a second language (the ES group, n =  `r per_group %>% pull(English)`, and another group spoke Spanish as their first language and English as their second (the SE group, n = `r per_group %>% pull(Spanish)`). Participants were screened prior to taking part in the study such that their age of onset was greater than 12 years old. 
Then, in order to examine how the same groups of bilinguals categorize different L3s, the SE and ES group were then randomly assigned either Hungarian (in the ES group, n = `r m_all_3_hu %>% pull(English)` and in the SE group, n = 
`r m_all_3_hu %>% pull(Spanish)`) or French (in the ES group, n = `r m_all_3_fr %>% pull(English)` and in the SE group, n = `r m_all_3_fr %>% pull(Spanish)`). 
In order to assess whether participants were eligible to take part in the study, a screening process was carried out in prolific which involved the Language History Questionnaire [@li_language_2020], and a proficiency test in the participant's L2. 
Speakers who reported knowing a third language or who scored below 60% on the lexTALE proficiency test were deemed ineligible for the study, and were not permitted to continue. 

Additionally, participants who failed attention checks, chose the same answer for every trial, or who answered over 80% /p/ or /b/ in the stops trials were removed from the data analysis. 
The data of `r m_english_removed_stops %>% pull(n)` participants were removed for English stops, `r m_spanish_removed_stops %>% pull(n)` for Spanish stops, `r m_french_removed_stops %>% pull(n)` for French stops, and `r m_hungarian_removed_stops %>% pull(n)` for Hungarian. The same procedure was carried out for vowels and a total of 5 participants' data was removed under the same criteria. 

## Materials

The experiment was composed of 5 total blocks. 
An itemized version of the Language History Questionnaire [@li_language_2020] was given to participants and asked their first language, second language, and the age at which they began learning their L2. 
The measure of L2 proficiency, the lexTALE [@izura_lexical_2016; @lemhofer_lexical_2016], was given in Spanish to the ES group and English to the SE group. 
The lexTALE is a lexical decision task in which the participants are presented with either a word or a non-word, and they must decide whether the word presented exists in that language.
Both the Spanish and English versions are scored in the same manner, and allow for comparability of proficiency across groups. Both versions contain 60 words total, 40 of which are real words and 20 of which are non-words. A formula then calculates a score based on the percentage of real words correct plus the percentage of non-words correct divided by 2. Table 1 provides descriptive statistics which shows the mean lexTALE score per L1 group. As the table shows, the Spanish L1 group was more proficient in their L2 on average than the English L1 group. Specifically, the mean lexTALE score of the Spanish L1 group was `r spanlexmean`, (SD = `r spanlexsd`) and the mean score of the English L1 group was `r englexmean`, (SD = `r englexsd`). 
Figure 1 shows a graphical relationship of the distribution of proficiency scores per L1 group. A test of equivalence of independent samples revealed that the Spanish L1 group and the English L1 group were not equivalent (*t*(50.52) = 0.957, p = 0.828), with equivalence bounds of -5.227 and 5.227 (on a raw scale) and an alpha of 0.05. A null hypothesis test of the same samples suggests that the null hypothesis is also rejected (*t*(50.52) = 2.958, p = 0.0047, given an alpha of 0.05). Concretely, these results suggest that the difference in proficiency between groups was `r tost_prof$diff` percentage points (90% CI `r tost_prof$LL_CI_TOST`-`r tost_prof$UL_CI_TOST`) on the lexTALE vocabulary task, with the Spanish group scoring slightly higher.

```{r}

desc_prof = read.csv(here("data", "tidy", "desc_prof.csv"))

apa_table(
  desc_prof
  , caption = "Descriptive Statistics of proficiency per L1 group"
  , escape = TRUE
)

```

```{r, fig.retina=2, fig.cap="Add figure caption here."}

desc_prof_plot = read.csv(here("data", "tidy", "desc_prof_plot.csv"))

desc_prof_plot %>% 
  ggplot(., aes(x = l1, y = score)) + 
  geom_point(alpha = .35, color = "seagreen") + 
  stat_summary(fun.data = mean_sdl, geom = "pointrange", pch = 21, 
               fill = "white", position = position_nudge(x = 0.1), size = 1, 
               fun.args = list(mult = 1)) + 
  xlab("L1 Group") + 
  ylab("lexTALE Score") + 
  ggtitle("Figure 1: Proficiency scores as a function of L1 group") + 
  theme_grey(base_family = "Times", base_size = 10)

# Slight adj to plot so that the points are visible
# I would use the fig.height and fig.width chunk options to make the plot smaller
```

Following these tasks, participants then took a series of three two-alternative forced choice tasks (2afc). 
The tasks were programmed in Psychopy [@peirce_psychopy2_2019] and were hosted on Pavlovia.org. 
Participants received written instructions prior to taking part in each phase of the experiment. The instructions for the "English" task were in English, the instructions for the "Spanish" task were in Spanish, and the instructions for the L3 tasks were bilingual (both Spanish and English). In each block, the instructions on profilic were in the language mode that the participants were told that they would hear, and the instructions within the experiment itself were also language specific, with the exception of all L3 prolific instructions being in English. These instructions read "Final step! You have been assigned to the Hungarian/French group.
The task is the same as the first two sessions, but this time, the words are from the Hungarian/French language." In the L3 experimental session, written instructions were more scarce, since the participants had already completed the same procedure twice, but were always presented in both English and Spanish at the same time. Instructions appeared at the beginning of the experiment, after the initial practice trials, and between the stop and vowel sessions, and for the attention checks. In each case, and English instruction was given and its Spanish translation appeared either to the right of the English instruction or immediately below it. 
As in previous studies, the instructions contained no instances of the experimental stimuli, in this case "pafri", "bafri, "ifri, or "ufri", or any of their language specific forms appeared during the instructions. 

## Stimuli 

The 2afc tested the perception of the same two continua (a /p/-/b/ VOT continuum and a /i/-/u/ F2 continuum) in three different language modes: English, Spanish and the L3 (either Hungarian or French). The VOT continuum has been used in previous studies [@lozano-arguelles_conceptually_2020; @gonzales_how_2019] and were created from a recording by a female Spanish-English bilingual. The original VOT continuum contained the word "pafri" and "bafri". The recording was resynthesized and ranged in VOT from -35ms to 35ms, excluding a 0 VOT step. In the present study, three additional steps were added on the positive end of the scale in order to increase the chances of /p/ responses at the positive extreme of the continuum, such that steps of 40ms, 45ms, and 50ms were included and made a total of 17 total steps. The extra steps were created manually in PRAAT by extracting 5ms of the voiceless part of the previous segments and adding these portions on to the beginning of the 35ms stimulus up to three times.
Just as in @lozano-arguelles_conceptually_2020, the present study conceptually cued participants, and did not utilize the language specific rhotics used in @gonzales_bafri_2013 and @casillas_perceptual_2018. 

The vowel continua were created using productions of the pseudowords "ifri" and "ufri" by a male American English speaker. These natural productions were then resynthesized using a PRAAT script which manipulated the second formant while holding the other formants constant. 
Just as was done with the stop continuum, only the first syllable together with [f] were produced in each stimulus to avoid perceptual cueing. As a result, the stimuli, ranging from "if" to "uf" varied only by the second formant (F2: a correlate for frontedness in the vowel space) in a total of 11 steps. 

During the 2afc task, participants saw both the words *pafri* and *bafri* or language specific versions of *ifri* and *ufri* on the screen and just the first syllable of these words in which the initial consonant was a member of the continuum ("paf" or "baf" for vowels, "if" or "uf" for vowels). 
Each step of the continuum was played 5 times total, leading to a total of 85 trials in the VOT continuum per participant and 55 total trials in the vowel continuum. 
Each trial was drawn randomly from the continuum in a single block, so no participant heard the same stimulus order. 
Due to language specific orthographic differences, the non-words on screen were distinct orthographically in each language, with the exception of French and Spanish.
In Spanish and French, the participants saw *ifri* and *ufri* on the screen, in English they saw *eefri* and *oofri*, and in Hungarian, they saw *ifrelo* and *ufrelo*. This spelling difference was necessary because of the orthographic conventions in the realizations of similar sounds between English and Spanish. For example, the Spanish grapheme "u" and the English graphemes "oo" are similar in their acoustic realizations. On the other hand, the choice of spelling in the French and Hungarian stimuli were made to reflect that high degree of lexical overlap between Spanish and French, and to reinforce the ambiguous relationship between Hungraian and English or Spanish.
 

## Procedure

All participants completed a total of five tasks, the Language History Questionnaire, the lexTALE proficiency test in their L2, and three 2afc tasks. First, the participants completed the Language History Questionnaire online [@li_language_2020], followed by the lexTALE proficiency test. 
Following these two tasks used for screening, only those participants who did have knowledge of a third language and scored at least 60% on the lexTALE were deemed eligible and continued the experiment. This cutoff was chosen to focus the analysis on more proficient bilinguals, rather than demonstrate that increasing proficiency is correlated to a larger double phonemic boundary effect size, as has been found in previous studies [@casillas_perceptual_2018; @garcia-sierra_testing_2009; @gonzales_bafri_2013].

Following these screening tasks, and in order to control for language mode, the participants completed the 2afc tasks in a semi-longitudinal design. 
Participants were invited to take part in the second portion of the experiment at least 30 minutes after their completion of the previous step. In many cases, however, participants completed different blocks of the experiment hours or days apart. 
All participants began with the "English" task, followed by "Spanish", and finished with either "French" or "Hungarian". Unfortunately, the lack of counterbalancing across participants is a limitation in the design of the current study.
In the English and Spanish tasks, the instructions were given in that respective language, and it was explained that the participants were going to hear rare words in the language in question (English, Spanish or the L3). 
In the L3 session, the same instructions were given in both English and Spanish simultaneously and on the same screen in order to avoid biasing the participants to either English or Spanish mode. 

## Statistical analyses 

In order to determine a crossover boundary for analysis in a general linear model, two logistic regression models were fit to each participant with their responses of ‘bafri’/‘pafri’, in the VOT continuum, and 'ifri/ufri', in the F2 continuum as a function of the standardized continuum step for each language. <!-- so 6 models to each participant, right?... 3 languages x 2 continua --> 
The continuum step standardization was done in order to allow for the comparability of the stop and vowel crossover boundaries, which had a distinct number of steps.
As a result, 6 crossover boundaries total were calculated per participant that provided which step in the continuum the probability of the participant choosing either "pafri" or "bafri" was 50%. These boundaries were used as continuous variables in subsequent general linear regression models and t-tests. All analyses were carried out in R [@R-base] and used either the stats package or the lmer package [@R-lme4]. In order to determine the independence of the crossover samples, paired t-tests were carried out in each language combination in both stops and vowels. A significant p-value on a paired t-test would indicate that there is non-zero difference between individual participants' crossover boundaries between the two languages.
