# Results 

## Descriptive statstics 

Tables 2 and 3 provide descriptive statistics of the crossover boundaries per group per language. The crossover boundary is the continuum step at which the probability of choosing both options (/p/ or /b/, or /i/ or /u/) was 50% and was calculated by fitting a logistic regression model to each participant in each language mode session. <!-- Well, that is half the story... you fit the mdoel to each individual becuase you need the intercept and slope to calculate the boundary --> 
The crossover boundaries were standardized to z-scores in order to allow for comparability across all language modes and over features (between stops and vowels). 

<!-- You need to add a sentence or two describing what the reader needs to take from these tables --> 

```{r}
apa_table(
  desc_vowels
  , caption = "Descriptive statistics of vowel crossovers by group"
  , escape = TRUE
)
```


```{r}
apa_table(
  desc_stops
  , caption = "Descriptive statistics of stops crossovers by group"
  , escape = TRUE
)
```

## Question 1: Was the double phonemic boundary effect replicted in stops and extended to vowels? 

In order to answer the research questions of the present study, it was vital to replicate previous findings of double phonemic boundary effect in Spanish-English bilinguals. Thus, an implicit research question in this study was whether Spanish-English bilinguals would show evidence of a double phonemic boundary in stops, as in previous studies, and, newly, whether this effect could be extended to vowels. Of course, without finding two language specific language mode-driven categorizations of English and Spanish, conclusions cannot be drawn as to whether bilinguals approach a third language using their first or second language perceptual routines. The double phonemic boundaries were assessed per language using a series of paired t-tests. The results showed that the English L1 group showed a double phonemic boundary effect in stops (`r el1ses$statistic`), but not in vowels (`r el1ves$statistic`). The Spanish L1 group showed evidence of the double phonemic boundary in both stops (`r sl1ses$statistic`) and vowels (`r sl1ves$statistic`). Figure 2 shows a graphical representations of the categorizations of vowels (figure 2) and stops (figure 3) in both languages in both L1 groups. 

```{r}
all_data = read.csv(here("data", "tidy", "all_2afc_long.csv"))

all_data %>% 
  filter(exp == "vowels", !(participant %in% vowels_remove)) %>% 
  filter(language == "english"| language == "spanish") %>% 
  ggplot(., aes(x = step_std, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 2: Spanish/English vowel boundaries per L1") + 
  theme_grey(base_family = "Times", base_size = 10)

# Same as above... adjust the plot so the font is a similar size and the 
# prose in the manuscript. It looks funny when it is this long. 
```

```{r}


all_data = read.csv(here("data", "tidy", "all_2afc_long.csv"))

all_data %>% 
  filter(exp == "stops", !(participant %in% stops_remove)) %>% 
  filter(language == "english"| language == "spanish") %>% 
  ggplot(., aes(x = step_std, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 3: Spanish/English Stop boundaries per L1") + 
  theme_grey(base_family = "Times", base_size = 10)

# Same as previous

```

## RQ2: which language-specific phonemic boundaries will participants perceive when they are introduced to the same VOT/F2 continuum in a “third language?” 

Due to the results of the initial series of paired t-tests, only the results of the stops in both L1 groups and the vowels in the Spanish groups were useful in answering this question, since English vowels violated the necessary assumption that all groups would show double phonemic boundary effects. A series of three paired t-tests are reported per each of the four participant groups. In each group, the paired t-tests serve to provide evidence that language categorizations are independent samples (if the p-value falls below .05) or not independent samples (if the p-value is greater that .05). 
The three t-tests in each group test the independence of their L1 and L2 categorization (in all 4 groups), the categorization of their L1 and their L3 (English or Spanish with Hungarian or French), and their L2 and their L3 (also English or Spanish with Hungarian or French). By having all three of the t-tests together, it can more easily be argued that one specific language has greater influence on L3 categorization if, first the Spanish and English paired t-test is significant, and one of the two remaining t-tests is significant and the other is not (i.e. if the L2 to L3 t-test is significant but the L1 to L3 t-test is not, that is taken as evidence that the L2 and the L3 are independent samples, but that the L1 and the L3 are not).

### Stops English L1 French L3 group  

The English L1 group appeared to use their L1 boundary in all three continua in both L1 groups. English L1 vowels did not show a group wide L1-L2 double phonemic boundary effect, and subsequent analyses in the English L1 groups focuses on stops.Despite the group-wide differences in the English L1 group, double phonemic boundary effects were not found in the individual English L1 groups. In stops, paired t-tests found non-significant results, interpreted as similar categorizations in English-Spanish categorization (`r el1fess$statistic`), Spanish-French categorization (L2 to L3) (`r el1ssf$statistic`) and English-French categorization (L1 to L3) (`r el1sef$statistic`). 

### Stops English L1 Hungarian L3 group

The English-Hungarian stops were found to be distinct. English-Hungarian categorization (L1 to L3) (`r el1seh$statistic`), but not the English-Spanish categorization (`r el1hess$statistic`) nor Spanish-Hungarian categorization (L2 to L3) (`r el1ssh$statistic`). 

### Stops Spanish L1 French L3 group

No categorization differences were found in Spanish L1 speakers assigned to the "French" group in stops. There was no significant English-Spanish categorization difference (`r sl1fess$statistic`), no significant English-French categorizations difference (`r sl1sef$statistic`) and no significant Spanish-French categorization difference (`r sl1ssf$statistic`).

### Vowels Spanish L1 French L3 group

Evidence of a double phonemic Spanish-English boundary was not found in this group (`r sl1fes$statistic`). Likewise, Spanish-French categorizations had no evidence of being distinct (`r sl1vsf$statistic`). There was, however, evidence of a distinction in English-French categorizations (`r sl1vef$statistic`). The lack of evidence for the double phonemic boundary effect in this group alone does not allow for definitive evidence that can inform which language boundary this group used in the perception of what they believed were French vowels.


### Stops Spanish L1 Hungarian L3 group

In this group, there was evidence that there was a English-Spanish categorization difference (`r sl1hess$statistic`), implying that this group showed evidence of the double phonemic boundary. The L2 and L3 English-Hungarian categorizations were also found to be categorized differently (`r sl1seh$statistic`), whereas the L1 and the L3 Spanish-Hungarian categorizations (`r sl1ssh$statistic`) were not categorized differently. 
These results suggest that the Spanish L1 group used their L1 boundaries to categorize what they believed were Hungarian stops.

### Vowels Spanish L1 Hungarian L3 group

The same trend is present in vowels as in stops for the Spanish L1 Hungarian L3 group. A paired t-test examining the cross over boundaries of English and Spanish (`r sl1hes$statistic`) reveals a double phonemic boundary effect. There was evidence that there is a difference in L2-L3 English-Hungarian categorizations (`r sl1veh$statistic`), but another paired t-test suggests that L1-L3 Spanish-Hungarian (`r sl1vsh$statistic`) are not categorized distinctly. Like stops, taken together, these results suggest that the Spanish L1 group used their L1 boundaries to categorize what they believed were Hungarian vowels. 


## RQ3: Did the L3 groups categorize the continua differently?

T-tests of independent samples revealed that L3 groups did not categorize their respective continua differently if they were assigned to the French group or the Hungarian group. The L3 French group with Spanish L1 and the L3 Hungarian group with a Spanish L1 did not categorize their respective L3s differently in vowels (`r sl1vfh$statistic`) or stops (`r sl1sfh$statistic`). The same evidence was found the the English L1 groups; they categorized L3 sounds similarly, whether they were in the L3 French group or Hungarian group in both vowels (`r el1vfh$statistic`) and stops (`r el1sfh$statistic`). <!-- So, they used the L1 boundary, right --> 
<!-- Not for these, these I used non-paired t-tests to see whether Hungarian and French L3 groups categorized their L3 differently -->
 

## Post-hoc analysis 

Based on the findings that the suggested that L3 continua were not categorized differently by the same L1 group, a post-hoc series of paired tests of equivalence were carried out in which the L3 groups were pooled together. 
First, tests of equivalence were carried out to determine whether L3 groups categorized the continua equivalently whether they believed the were hearing Hungarian or French.
Importantly, the same participants did not hear both French and Hungarian, but were randomly assigned to one of them. As a result, a non-paired test of equivalence was carried out. Cohen's D was set to .5 in these tests, which has been reported as a medium effect size in the literature [@cohen_statistical_2013]. 
All tests were carried out in R [@R-base], using the `TOSTER` function [@toster].

Figure 4 shows a graphical summary of three non-paired tests of equivalence. The first was a comparison between categorization of vowels between the French L3 and Hungarian L3 groups with Spanish as their L1. Surprisingly, the equivalence test was did not quite show that the categorizations of French and Hungarian vowels by the Spanish L1 groups were equivalent (t(31.25) = -1.510, p = 0.0705, given equivalence bounds of -0.0266 and 0.0266 (on a raw scale) and an alpha of 0.05). 
It is likely that, it this case, the lack of equivalence found is due to a low sample size. 

The second test of equivalence measured the categorizations of stops between the French and Hungarian L3 groups who speak L1 Spanish. These results are similar to those reported in vowels in that they do not provide enough evidence that the L3 groups categorized what they believed to be L3 sounds the same (t(37.81) = -1.118, p = 0.135, given equivalence bounds of -0.248 and 0.248 (on a raw scale) and an alpha of 0.05). 
Again, this is suggested to be due to low sample size. 

The third test of equivalence measured the categorizations of stops between the French and Hungarian L3 groups who speak L1 English. A similar trend was observed in this case, in which equivalence was not detected, likely due to low sample size (t(10.89) = -0.829, p = 0.212, given equivalence bounds of -0.242 and 0.242 (on a raw scale) and an alpha of 0.05.).

```{r}
all_tost_l3groups %>% 
  ggplot(aes(x = diff, y = combination)) + geom_point() + 
  geom_linerange(aes(xmin = LL_CI_TOST, xmax = UL_CI_TOST)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle("Figure 4: Categoriztions between L3 groups") + xlab("Difference") + 
  theme_grey(base_family = "Times", base_size = 10)

# Same as previous
```

Inspired by the idea that low sample size may be the cause of the lack of equivalence in the findings, an additional series of paired equivalence tests were carried out in Spanish L1 vowels, Spanish L1 stops, and English L1 stops. English L1 vowels were omitted due to the lack of any recognizable trend towards a meaningful difference in the initial analysis.

### Spanish L1 group pooled stops 

Firstly, this group showed evidence of the double phonemic boundary effect when they were pooled (`r pooled_stops_sl1_se$statistic`).
Next, of interest in this group was whether the Spanish L1 group categorized the L3 stop continua using their L1 boundaries. The test of equivalence suggest that this is the case (t(45) = -3.094, p = 0.00169). 
On the other hand, English and the L3 were not found to be equivalent (t(45) = -1.103, p = 0.138), and nor were Spanish and English (t(45) = -0.770, p = 0.223).
Figures 5 is a graphical representation of the crossover boundaries of stops in the pooled data of both Spanish L1 groups.
Figure 6 shows the effect sizes derived from each of the three tests of equivalence. 
As can be seen in figure 6, the smallest effect size is the difference between the L1 and the L3, which is rather close to zero. 

```{r}


all_data = read.csv(here("data", "tidy", "all_2afc_long.csv"))

all_data %>% 
  filter(exp == "stops", !(participant %in% stops_remove)) %>% 
  filter(l1 == "Spanish") %>% 
  mutate(language = ifelse(as.character(language) == "french", "l3", as.character(language))) %>% 
  mutate(language = ifelse(as.character(language) == "hungarian", "l3", as.character(language))) %>% 
  ggplot(., aes(x = step_std, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /p/") + xlab("VOT") + 
  ggtitle("Figure 5: Spanish L1 group pooled stops") + 
  theme_grey(base_family = "Times", base_size = 10)


all_tost_spanl1_stops %>% 
  ggplot(aes(x = diff, y = combination)) + geom_point() + 
  geom_linerange(aes(xmin = LL_CI_TOST, xmax = UL_CI_TOST)) +
  geom_vline(xintercept = all_tost_spanl1_stops$low_eqbound_dz, linetype = "dashed") + 
  geom_vline(xintercept = all_tost_spanl1_stops$high_eqbound_dz, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = .5) +
  ggtitle("Figure 6: Spanish L1 stops TOST") + 
  theme_grey(base_family = "Times", base_size = 10)

# same as previous

```

### Spanish L1 group pooled vowels 

A similar results was found in the Spanish L1 group in vowels. First, there was evidence of a double phonemic boundary (`r pooled__vowel_sl1_se$statistic`) between English and Spanish. 
Next, there is evidence that the L1 and the L3 were categorized equivalently (t(45) = -2.949, p = 0.00252), where the L2 and L3 (t(45) = 0.198, p = 0.578) and the L1 and the L2 (t(45) = 0.536, p = 0.703) were not. 
Just as in stops, Figure 7 is a graphical representation of the crossover boundaries of vowels in the pooled data of both Spanish L1 groups.
Figure 8 shows the effect sizes derived from each of the three tests of equivalence. 
Again, similarly to stops, in figure 8, the smallest effect size is the difference between the L1 and the L3, which is rather close to zero. 


```{r}

all_data %>% 
  filter(exp == "vowels", !(participant %in% vowels_remove)) %>% 
  filter(l1 == "Spanish") %>% 
  mutate(language = ifelse(as.character(language) == "french", "l3", as.character(language))) %>% 
  mutate(language = ifelse(as.character(language) == "hungarian", "l3", as.character(language))) %>% 
  ggplot(., aes(x = step_std, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /u/") + xlab("Continuum Step") + 
  ggtitle("Figure 7: Spanish L1 group pooled vowels") + 
  theme_grey(base_family = "Times", base_size = 10)


all_tost_spanl1_vowels %>% 
  ggplot(aes(x = diff, y = combination)) + geom_point() + 
  geom_linerange(aes(xmin = LL_CI_TOST, xmax = UL_CI_TOST)) + 
  geom_vline(xintercept = all_tost_spanl1_vowels$low_eqbound_dz, linetype = "dashed") + 
  geom_vline(xintercept = all_tost_spanl1_vowels$high_eqbound_dz, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = .5) +
  ggtitle("Figure 8: Spanish L1 Vowels TOST") + 
  theme_grey(base_family = "Times", base_size = 10)

# same as previous

```

### English L1 group pooled stops 

Overall, clear conclusions cannot be drawn regarding the English L1 group's categorization of stops. 
In the fist analysis, only the stops showed evidence of a double phonemic boundary modulated by language mode, however, this Spanish-English double phonemic boundary effect does not quite persist in those who completed all phases of the experiment (`r pooled_stops_el1_se$statistic`).

The data also suggested that this group showed evidence of their L2 influencing L3 comprehension, rather than their L1. The equivalence test between L2 and L3 was the closest to equivalent, t(22) = 1.441, p = 0.0818), but did not provide sufficient evidence to classify the data as evidence for an L2 status effect. 
Importantly, this groups L1 and L3 categorizations were not equivalent (t(22) = 0.427, p = 0.663), and neither were their L1 and L2 categorizations (t(22) = -0.527, p = 0.302)
Just as in the previously reported tests of equivalence, figures 9 shows the graphical representation of the crossover boundaries of stops in the pooled data of both English L1 groups.
Figure 10 shows the relative effect sizes of each language combination. 
Taken together, it is difficult to draw clear conclusions in regard to the English L1 group. 

```{r}

all_data %>% 
  filter(exp == "stops", !(participant %in% stops_remove)) %>% 
  filter(l1 == "English") %>% 
  mutate(language = ifelse(as.character(language) == "french", "l3", as.character(language))) %>% 
  mutate(language = ifelse(as.character(language) == "hungarian", "l3", as.character(language))) %>% 
  ggplot(., aes(x = step_std, y = response, color = language)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_grid(. ~ l1) + ylab("Proportion /p/") + xlab("VOT") + 
  ggtitle("Figure 9: English L1 group pooled stops") + 
  theme_grey(base_family = "Times", base_size = 10)


all_tost_engl1_stops %>% 
  ggplot(aes(x = diff, y = combination)) + geom_point() + 
  geom_linerange(aes(xmin = LL_CI_TOST, xmax = UL_CI_TOST)) +
  geom_vline(xintercept = all_tost_engl1_stops$low_eqbound_dz, linetype = "dashed") + 
  geom_vline(xintercept = all_tost_engl1_stops$high_eqbound_dz, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = .5) +
  ggtitle("Figure 10: English L1 stops") + 
  theme_grey(base_family = "Times", base_size = 10)

# same as previous

```



